\documentclass[a4paper, 10pt,twoside]{report}
\usepackage[utf8x]{inputenc}

% english language
\usepackage[UKenglish]{babel}

% ps tricks
\usepackage[usenames,dvipsnames]{pstricks}

% about title
\usepackage{titlesec}

% margins
\topmargin 0cm
\headsep 1cm
\headheight 0.6cm
\textwidth 16.6cm
\textheight 21.8cm
\evensidemargin -0.5cm
\oddsidemargin -0.5cm

% mathematics
\usepackage{amsmath}
\usepackage{amsfonts}


% definition
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Proof.}

\theoremstyle{definition}
\newtheorem{defn}{Definition}[section]

\theoremstyle{remark}
\newtheorem{rmk}{Remark}

% test format
\usepackage{verbatim}
\usepackage{color}
\usepackage{eurosym}

% symbols
\usepackage{textcomp, latexsym, amssymb, gensymb, placeins}
\usepackage[mathscr]{euscript}
\usepackage{amsmath}

% figures
\usepackage{epsfig}
\usepackage{graphics, epsfig, color, colortbl, supertabular}
\usepackage{subfigure}
\usepackage{epstopdf}

% tables
\usepackage{multirow, ctable, rotating, lscape}
\usepackage{array}
\usepackage{booktabs}
\usepackage{supertabular}
\usepackage{tabularx}
\usepackage{graphicx}

% other packages
\usepackage{hyperref}
\def\thesection{\arabic{section}}
\def\theequation{\arabic{equation}}

\addtocounter{section}{1}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\setcounter{section}{0}

\usepackage{float}
\usepackage{stmaryrd}


% tikz
\usepackage{tikz}
\usetikzlibrary{mindmap,trees}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,fit,positioning,%
  shapes.symbols,chains}
\usetikzlibrary{calc,trees,positioning,arrows,chains,shapes.geometric,%
  decorations.pathreplacing,decorations.pathmorphing,shapes,%
  matrix,shapes.symbols}


% acronyms
\usepackage{acronym}


% URL
\usepackage{url}

\usepackage{fancyhdr}
\usepackage{etex}
\usepackage{epstopdf}

% X Quartapelle
% \renewcommand{\theequation}{\thechapter.\arabic{equation}}
% \newcommand{\clearemptydoublepage}{\newpage{\pagestyle{empty}
% \cleardoublepage}}
\renewcommand{\headrulewidth}{0.3pt}
\renewcommand{\footrulewidth}{0.3pt}
\newcommand{\virgolette}[1]{``#1''}

% Mathematical Operator
\newcommand{\WE}{\mathcal{W}}
\newcommand{\W}{$\mathcal{W}$}
\newcommand{\LE}{\mathcal{L}}
\newcommand{\Lne}{$\mathcal{L}$}
\newcommand{\delOp}{\widehat{\boldsymbol{\nabla}}}
\newcommand{\diver}[1]{\text{Div}\big(#1\big)}
\newcommand{\invariants}{$\big(I_{\C},II_{\C},III_{\C}\big)$}

% Eucledian spaces
\newcommand{\Real}{$\mathbb{R}^3$}
\newcommand{\RN}{\mathbb{R}^{N_h}}

% Lagrangian and Eulerian displacements
\newcommand{\displL}{\widehat{\underline{\eta}}}
\newcommand{\displE}{\underline{\eta}}

% Discrete displacement, velocity and acceleration
% The fist command is to reuse the same already existing formula
\newcommand{\Spost}{\widehat{\underline{\Lambda}}}
\newcommand{\GradSpost}{\delOp\delta\Spost}
\newcommand{\vectL}{\widehat{\underline{\Lambda}}}
\newcommand{\velVect}{\widehat{\underline{W}}}
\newcommand{\acc}{\widehat{\underline{A}}}

% Stiffness Vectors, discretized system
\newcommand{\StiffVect}{\underline{K}(\vectL)}
\newcommand{\StiffDiscr}[1]{\underline{K}(\vectL^{#1})}

% Lagrangian velocity
\newcommand{\velL}{\widehat{\underline{v}}}
\newcommand{\velE}{\underline{v}} %Test function (velocity)
\newcommand{\velwE}{\widehat{\underline{w}}}

% Programming names
\newcommand{\SSol}{\textit{Structural Solver}}
\newcommand{\SSolNC}{Structural Solver}
\newcommand{\LV}{\texttt{LifeV}}
\newcommand{\tPC}[1]{\texttt{#1}}
\newcommand{\AES}{AssemblyElementalStructure}
\newcommand{\tSS}{\textit{test\_structuralsolver}}

% Reference and Current Configuration
\newcommand{\RefCon}{$\mathcal{B}_0$}
\newcommand{\CurCon}{$\mathcal{B}(t)$}
\newcommand{\RefConE}{\mathcal{B}_0}
\newcommand{\CurConE}{\mathcal{B}(t)}

% Position in the reference and current configuration
\newcommand{\posE}{\underline{X}}
\newcommand{\nposE}{\underline{x}}
\newcommand{\pos}{$\underline{X}$}
\newcommand{\npos}{$\underline{x}$}

% Tensors and matrices
\newcommand{\F}{\textbf{F}}
\newcommand{\cofF}{\text{Cof}\textbf{F}}
\newcommand{\T}{\textbf{T}}
\newcommand{\D}{\textbf{D}}
\newcommand{\C}{\textbf{C}}
\newcommand{\Cbar}{\bar{\textbf{C}}}
\newcommand{\I}{\textbf{I}}
\newcommand{\mass}{\textbf{M}}
\newcommand{\Piola}{\textbf{P}}
\newcommand{\Et}{\textbf{E}}

% Expression for the Exponential-Material
\newcommand{\term}{e^{\gamma\left(I_{\C}-3\right)}}

\newtheorem{weak}{Weak-formulation}[chapter]
\newtheorem{problem}{Problem}[chapter]

% Title Page
\title{Structural Solver framework in \textit{LifeV}:
Description and Usage}
\author{Paolo Tricerri, Gianmarco Mengaldo}


\begin{document}

\maketitle


\section*{Summary} The aim of the present document is to describe the
\SSol framework in the open source finite element library LifeV. This
framework has been developed to have a unique solver for structural
dynamics problems (described by the equation of motion) letting the
users plug different constitutive laws into the conservation
equation of linear momentum in a flexible way.\\ In this report, we
aim at:
\begin{itemize}
\item presenting the architecture of the StructuralSolver;
\item giving a detailed description of the constitutive laws
  implemented;
\item showing the usage of \textit{test\_structuralsolver} that has
  been developed and, to conclude, presenting the validation of the
  implemented laws, with respect to analytical and semianalytical
  solutions;
\end{itemize} This document is divided in different sections. The
first one briefly introduces the continuum mechanics theory behind the
formulation of the conservation equation of linear momentum. After
that, the mathematical formulation of the constitutive laws available
in the library is presented. The second section deals with the
discretization of the continuous differential problem and its
formulation in the finite element method. Moreover, the linearization
of the nonlinear algebraic system at each time step is shown. The
third part describes the main features of the \SSol framework and the
implementation of the constitutive laws defined in the first
section. The last section describes the
\textit{test\_structuralsolver} for a user who wants to solve
structural dynamics problems. Starting from the case solved in the
test, the section deals with the validation of the different material
laws, with respect to an analytical and semianalytical solution for a
traction test on a cube.\\

\section*{Acknowledgements} The authors would like to thank
Dr. Mariarita Deluca for her help during the development and debug of
the code and for implementing a first version of the structural models
in the serial version of the library.

\tableofcontents
\newpage


\input{FirstSection.tex}
\input{SecondSection.tex}
\input{ThirdSection.tex}
\input{FourthSection.tex}

\bibliographystyle{plain}
\bibliography{StructuralSolver}
\addcontentsline{toc}{section}{Bibliography}

\end{document}
