%
%
% SUMMARY:
% USAGE:
%
% AUTHOR:       Gilles Fourestey
% ORG:          EPFL
% E-MAIL:       foureste@iacspc.epfl.ch
%
% ORIG-DATE:  4-Feb-09 at 14:11:50
% LAST-MOD:  5-Feb-09 at 18:11:24 by Gilles Fourestey
%
% DESCRIPTION:
% DESCRIP-END.


Now that we have decribed a simple stationary problem, let's have a look at the evolutionary
case. In this example, we will consider the same domain, but this time we will solve the
Navier-Stokes problem \ixn{Navier-Stokes Problem}. Starting from the Oseen problem (\ref{eqn-oseen})

\begin{equation*}
\left\{
\begin{array}{rl}
\displaystyle \alpha u + \beta \cdot \nabla u - \nu \Delta u+
\nabla p & = f \\
\displaystyle \nabla \cdot u & = 0
\end{array}
\right.
\end{equation*}

We want to solve the instationary Navier-Stokes problem :

\begin{equation*} \label{eqn-navierstokes}
\left\{
\begin{array}{rl}
\displaystyle \frac{\partial u}{\partial t} + u \cdot \nabla u - \nu \Delta u+
\nabla p & = f \\
\displaystyle \nabla \cdot u & = 0
\end{array}
\right.
\end{equation*}

which can be written, using a time semi-discretisation of the time partial derivative :

\begin{equation*} \label{eqn-nsnl}
\left\{
\begin{array}{rl}
\displaystyle \alpha u^{n+1} + u^{n+1} \cdot \nabla u^{n+1} - \nu \Delta u^{n+1}+
\nabla p^{n+1} & = f^n  \\
\displaystyle \nabla \cdot u & = 0  \\
\end{array}
\right.
\end{equation*}

Where $alpha$ is a constant which depends on the time discritization order, $\Delta t$ is the time step, $u^{n+1}$ and $p^{n+1}$ are the velocity and the pressure at the time $t^n = n\Delta t$.
Note that $f^n$ now contains terms resulting from the time discretization.
Using a linearaztion $\beta^n$ of (\ref{eqn-nsnl}) around $u^{n+1}$, we get the full semi-discretized linear Navier-Stokes
equations :

\begin{equation*} \label{eqn-ns}
\left\{
\begin{array}{rl}
\displaystyle \alpha u^{n+1} + \beta^n \cdot \nabla u^{n+1} - \nu \Delta u^{n+1}+
\nabla p^{n+1} & = f^n  \\
\displaystyle \nabla \cdot u & = 0  \\
\end{array}
\right.
\end{equation*}

Solving (\ref{eqn-ns}) using the framework we used for the stationary driven cavity is not difficult :
instead of setting $alpha$, $beta$ and $f^n$ to zero, we give them their proper values. For instance,
is the consider the first order time discretization, we get :

\begin{equation*} \label{eqn-nso1}
\left\{
\begin{array}{r;}
\displaystyle \frac {u^{n+1}}{\Delta t} + u^n \cdot \nabla u^{n+1} - \nu \Delta u^{n+1}+
\nabla p^{n+1} & = \displaystyle \frac{u^n}{\Delta t}  \\
\displaystyle \nabla \cdot u & = 0  \\
\end{array}
\right.
\end{equation*}

The second order time discretization is :

\begin{equation*} \label{eqn-nso2}
\left\{
\begin{array}{rl}
\displaystyle \frac {3u^{n+1}}{2\Delta t} + u^n \cdot \nabla u^{n+1} - \nu \Delta u^{n+1}+
\nabla p^{n+1} & = \displaystyle \frac{2u^n}{\Delta t} - \frac{u^{n-1}}{2\Delta t}  \\
\displaystyle \nabla \cdot u & = 0  \\
\end{array}
\right.
\end{equation*}

Let's have a look at the code in \verb!testsuite/test_cavity/main.cpp!.
Until the first \verb!fluid.iterate()!, the code is the same as the one used to compute the Stokes problem.
Now, we want to use this Stokes solution to initialize the our instationary Navier-Stokes problem, and be able
to store a history of previous solution in order to be able to compute the discretized time derivative.
This can be done by using the following object :
\begin{verbatim}
    // bdf initialization with the stokes problem solution
    BdfTNS<vector_type> bdf(dataNavierStokes.order_bdf());
\end{verbatim}
The Backward Differentiation Formula templated class \verb!BdfTNS! \ixn{Backward Differentiation Formula} will store the previous solution $u^n$, $u^{n-1}$ ... so that
they will always be accessible when we need them. All we have to do is to construct this class
with the correct time discretization order given in the data file by the variable \verb!fluid/discretization/bdf_order!
so that the storing vector will be resized correctly, and intialize it with our previously computed Stokes problem solution :

\begin{verbatim}
   bdf.bdf_u().initialize_unk( fluid.solution() );
\end{verbatim}

We are now ready to enter the time loop :

\begin{verbatim}

    Real dt     = dataNavierStokes.timestep();
    Real t0     = dataNavierStokes.inittime();
    Real tFinal = dataNavierStokes.endtime ();

    int iter = 1;

    for ( Real time = t0 + dt ; time <= tFinal + dt/2.; time += dt, iter++)
    {
        // inside the time loop, it's really like the initialization procedure,
        // exept that we now have an advection velocity, rhs and the mass matrix
        dataNavierStokes.setTime(time);

        if (verbose)
        {
            std::cout << std::endl;
            std::cout << "We are now at time " << dataNavierStokes.time()
                      << " s. " << std::endl;
            std::cout << std::endl;
        }

        chrono.start();

        // alpha coefficient for the mass matrix
        double alpha = bdf.bdf_u().coeff_der( 0 ) / dataNavierStokes.timestep();

        // extrapolation of the advection term
        beta = bdf.bdf_u().extrap();

        // rhs  part of the time-derivative
        rhs  = fluid.matrMass()*bdf.bdf_u().time_der( dataNavierStokes.timestep() );

        // the we update the Oseen system
        fluid.updateSystem( alpha, beta, rhs );

        // and we solve it
        fluid.iterate( bcH );

        // shifting the previous solutions
        bdf.bdf_u().shift_right( fluid.solution() );

        // and we postprocess
        *velAndPressure = fluid.solution();
        ensight.postProcess( time );

        // a barrier to make sure everyone is here, and we start again
        MPI_Barrier(MPI_COMM_WORLD);

        chrono.stop();
        if (verbose) std::cout << "Total iteration time "
                               << chrono.diff()
                               << " s." << std::endl;
    }

\end{verbatim}

The time step \verb!dt!, the initial time \verb!t0! and the final simulation time \verb!tFinal!
are found in the data file variables respectively name \verb!fluid/discretization/timestep!,
\verb!problem/Tstart! and \verb!fluid/physics/endtime!. As we can see, a time step can be describe
as a follow up of several intuitive calls :
\begin{itemize}
\item computation of $\alpha$ ( which should is constant in most cases ),
\item computation of $\beta$ using the \verb!Bdf! class,
\item computation of the Right Hand Side $rhs$,
\item update of the system using these three variables,
\item resolution of the linear system.
\end{itemize}
After the system is solve, we simply update all time-depend variables such as the storage
of the previous solutions, and we restart at the begining of the loop until all time steps are computed.


%
%%%%%%%%%%%%% Some Settings for emacs and auc-TeX
% Local Variables:
% TeX-master: t
% TeX-command-default: "PDFLaTeX"
% TeX-parse-self: t
% TeX-auto-save: t
% x-symbol-8bits: nil
% TeX-auto-regexp-list: TeX-auto-full-regexp-list
% eval: (ispell-change-dictionary "american")
% End:
%
